Deep Probabilistic Models for Camera Geo-Calibration

### 2018-08-31 chapters/introduction/introduction.tex
check a lot of things

Most recent geo-calibration algorithms are deterministic systems
\todo{citations}, 
which have some obvious drawbacks: 1) they can not model the inherent
uncertainties from images of ambiguous scenes, and 2) without a proper
modeling of the uncertainty, the subsequent process can yield overly
confident predictions. To address these problems, we propose to build
probabilistic models for camera geo-calibration.

We also show that learning to geo-calibrate a camera allows us to
learn to understand the scene.
As an essential part of scene understanding, how to extract image
features is an important problem being studied for decades.  In
recent years, learning based approaches for image feature extraction
like convolutional neural networks have drawn a lot of attention due
to the fact that they do not need expert knowledge about the target
data. However, most of these feature learning methods require large
quantity of manually labeled data sets.  In this thesis, we presented
alternative ways to learn high-level image features when the
manually labeled data is either insufficient or absent.

\section{Background}
For a better understanding of our work, we would like to introduce
some related computer vision concepts in this section before the
discussion of the main topics.

\subsection{Geo-Calibration of Outdoor Images}
Automatic geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet. Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}. Conceptually the geo-calibration
task includes: identifying the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the query image was captured ({\em time estimation}).

@@

\item \textbf{Pose Estimation:}
Pose estimation includes estimating the camera yaw, pitch, and roll
angles, $(\theta_{yaw}, \theta_{pitch}, \theta_{roll})$ from a
image. In the context of geo-calibration, the yaw angle
$\theta_{yaw}$ is referred to as the geographic orientation angle of
the camera.
%
Li et al.~\cite{li2012worldwide} exploit geo-registered 3D points
clouds to estimate camera pose.
Vo \etal~\cite{vo2016localizing} propose a geo-localization network
that can regress the geo-orientation angle of the camera.
Agarwal \etal~\cite{agarwal2015metric} keep track of the camera pose
changes by matching SIFT~\cite{lowe1999object} feature points detected
between the input image and Google Street View panoramas.
%
Horizon line detection is an essential step for some pose
estimation algorithms, because horizon lines are closely related to
camera pitch and roll angles.
%If the camera focal length is known, one can compute the camera pitch
%and roll angle from the horizon line.
Collins and Weiss~\cite{unitsphere1990} formulate horizon line
detection as a statistical estimation problem on the Gaussian Sphere,
which is similar to the geometry we use.  More recent work has
explored the use of dual space~\cite{alignment2014,dualspace2013}
representations. Among the clustering-based approaches, Xu et
al.~\cite{kitware2013} improve this pipeline by introducing a new
point-line consistency function that models errors in the line segment
extraction step.
\newline

\item \textbf{Geo-Localization:}
The task of geo-localization estimates the geographic location, $(lat,
lon)$, of the camera from the input image or video frames.
%
As an important step of geo-localization, extracting
location-dependent features from image data has drawn a great detail
of attention from the vision community~\cite{jacobs07geolocate,
jacobs11geolocate, jacobs08geoorient}. The common trend amongst these
methods is that they take advantage of a large dataset of
geo-referenced images. Hays and Efros~\cite{hays2008im2gps} use a
data-driven scene matching approach to localize a query image using a
large dataset of geo-tagged images.  Doersch
\etal~\cite{doersch2012what} extract location-dependent features that
capture the relative appearance differences of large cities.  Lin et
al.~\cite{lin2013cross} localize a ground-level image by learning the
relationship between pairs of ground and aerial images \footnote{In
the context of our thesis, we do not distinguish between satellite
imagery and aerial imagery} of the same location. Other techniques
focus on urban environments and infer location using local image
descriptors~\cite{schindler2008detecting,snavely2006photo}.  Many
other cues exist, such as the
skyline~\cite{baatz2012large,ramalingam2009geolocalization}, sky
appearance~\cite{lalonde2010sun,workman2014rainbow}, and
shadows~\cite{junejo2008estimating,wu2010geo}.
\newline

\item \textbf{Time Estimation:}
The goal of time estimation is to estimate the time when the input
image was captured. Depending on the application, the scale of a
time estimator outputs may range from year to hour.
%
Matzen and Snavely~\cite{matzen2014scene} predict
timestamps for photos by matching against a time-varying
reconstruction of a scene.  Hill~\cite{hill1994elephant} estimate the
time of the day by measuring the light intensity profiles captured by
cameras. Methods are proposed to date the
year book photos by analyzing human
appearance~\cite{salem2016face2year,ginosar2015century}. Lee
\etal~\cite{linking2015iccp} find visual paterns in the buildings,
relate them to certain time periods.
\newline

@@

In conclusion, most of the geo-calibration algorithms we mention above
relies on finding visual cues from the input images. The development
of these algorithms require strong human expertise about the data,
thus can only be applied in limited scenarios. Therefore, researchers
need new methodologies in order to make the development of
geo-calibration algorithms more efficient.

@@

These geo-calibration methods make the predictions directly out of
DNNs.  There is another big category of geo-calibration methods that
use aerial imagery as geo-registered database, which we will introduce
in the following section.

@@

The key for the success of these ground-to-aerial geo-calibration
methods is to learn descriptive features good for ground-to-aerial
imagery matching. In the following section, we will show how to use
geo-calibration information to learn useful features.

\subsection{Geo-Calibration for Feature Learning}

Manually engineering visual features is a common practice for many
computer vision tasks. One of the successful example of manually
engineered feature is the SIFT feature~\cite{lowe1999object}.
In recent years, automatic feature engineering has drawn a lot
of attention. Compared to the manual feature
engineering, automatic feature engineering approaches allow to learn
useful features without strong expert knowledge about the data. Among
these approaches, deep neural networks are wildly accepted as good
tools to learn high-level image representations.  The most commonly
used approach for deep feature learning is full supervision, which
usually requires large number of manual
annotations~\cite{yosinski2014transferable,zhou2016learning,wen2016discriminative}.
Unfortunately, such data is not always available for certain tasks or
needs to invest a huge amount of human labors to create.
In contrast, unlabeled data is often plentiful (\ie cellphone photos
in Instagram or Youtube videos).

@@

So far, we have explained four important related computer vision
concepts. In the following section, we are going to discuss our main
contributions in this thesis.

>>>>

Most recent geo-calibration algorithms are deterministic systems
\todo{citations}. 
Deterministic systems have some obvious drawbacks: 1) they can not
model the inherent uncertainties from images of ambiguous scenes, and
2) without a proper modeling of the uncertainty, the subsequent
process can yield overly confident predictions. To address these
problems, we propose to build probabilistic systems for camera
geo-calibration.

We also show that learning to geo-calibrate a camera allows us to
learn about the scene of the image.
As an essential part of scene understanding, how to extract image
features is an important problem being studied for decades.  In
recent years, learning based approaches for image feature extraction
like convolutional neural networks have drawn a lot of attention due
to the fact that they do not need expert knowledge about the target
data. However, most of these feature learning methods require large
quantity of manually labeled datasets.  In this thesis, we presented
alternative ways to learn image features when the
manually labeled data is either insufficient or absent.

\section{Background}
To give a better understanding of our work, we would like to introduce
some related computer vision concepts before we discuss the main topics.
We start with the definition of geo-calibration,
along with its literature. Next, we introduce the history of deep
learning methods for geo-calibration, and a big category of approaches
that use aerial imagery. In the end, we cover some related work
of weakly supervised learning techniques.

\subsection{Geo-Calibration of Outdoor Images}
Automatic geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet. Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}. Conceptually the geo-calibration
task includes identifying: the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the image was captured ({\em time estimation}).

@@

\item \textbf{Pose Estimation:}
Pose estimation is to estimate the camera yaw, pitch, and roll
angles, $(\theta_{yaw}, \theta_{pitch}, \theta_{roll})$ from a
image. In the context of geo-calibration, the yaw angle
$\theta_{yaw}$ is referred to as the geographic orientation angle of
the camera.
%
Li et al.~\cite{li2012worldwide} exploit geo-registered 3D points
clouds to estimate camera pose.
Vo \etal~\cite{vo2016localizing} propose a geo-localization network
that can regress the geo-orientation angle of the camera.
Agarwal \etal~\cite{agarwal2015metric} keep track of the camera pose
changes by matching SIFT~\cite{lowe1999object} feature points detected
between the input image and Google Street View panoramas.
%
Horizon line detection is an essential step for some pose
estimation algorithms, because horizon lines are closely related to
camera pitch and roll angles.
Collins and Weiss~\cite{unitsphere1990} formulate horizon line
detection as a statistical estimation problem on the Gaussian Sphere,
which is similar to the geometry we use.  More recent work has
explored the use of dual space~\cite{alignment2014,dualspace2013}
representations. Among the clustering-based approaches, Xu et
al.~\cite{kitware2013} improve this pipeline by introducing a new
point-line consistency function that models errors in the line segment
extraction step.
\newline

\item \textbf{Geo-Localization:}
The task of geo-localization estimates the geographic location, $(lat,
lon)$, of the camera given image/video inputs.
%
As an important step of geo-localization, extracting
location-dependent features from image data has drawn a great detail
of attention from the vision community~\cite{jacobs07geolocate,
jacobs11geolocate, jacobs08geoorient}. The common trend amongst these
methods is that they take advantage of a large dataset of
geo-referenced images. Hays and Efros~\cite{hays2008im2gps} use a
data-driven scene matching approach to localize a query image using a
large dataset of geo-tagged images.  Doersch
\etal~\cite{doersch2012what} extract location-dependent features that
capture the relative appearance differences of large cities.  Lin et
al.~\cite{lin2013cross} localize a ground-level image by learning the
relationship between pairs of ground and aerial images \footnote{In
the context of our thesis, we do not distinguish between satellite
imagery and aerial imagery} of the same location. Other techniques
focus on urban environments and infer location using local image
descriptors~\cite{schindler2008detecting,snavely2006photo}.  Many
other cues exist, such as the
skyline~\cite{baatz2012large,ramalingam2009geolocalization}, sky
appearance~\cite{lalonde2010sun,workman2014rainbow}, and
shadows~\cite{junejo2008estimating,wu2010geo}.
\newline

\item \textbf{Time Estimation:}
The goal of time estimation is to estimate the time when the 
image was captured. Depending on the application, the accuracy of
predictions range from hour to year.
%
Matzen and Snavely~\cite{matzen2014scene} predict
timestamps for photos by matching against a time-varying
reconstruction of a scene.  Hill~\cite{hill1994elephant} estimate the
time of the day by measuring the light intensity profiles captured by
cameras. Methods are proposed to date the
year book photos by analyzing human
appearance~\cite{salem2016face2year,ginosar2015century}. Lee
\etal~\cite{linking2015iccp} find visual paterns in the buildings,
relate them to certain time periods.
\newline

@@

In conclusion, most of the geo-calibration algorithms we mention above
relies on finding visual cues from input images. To extract useful
visual features majority of these algorithms require strong human
expertise about the data, thus can only be applied in limited
scenarios. Therefore, we contribute to developing geo-calibration
algorithms without need of strong human expertise about the data.

@@

All these geo-calibration methods share the common property that they
make the predictions directly out of DNNs. Besides these methods,
there exists another big category of methods that geo-calibrate
cameras using aerial imagery as geo-registered database. We will
discuss them in the following section.

@@

The key for the success of these methods is to learn good descriptive
features for ground-to-aerial imagery matching. In the following
section, we will introduce literature about feature learning techniques.

\subsection{Feature Learning with Weak Annotations}

Manually engineering visual features is a common practice for
computer vision tasks. One of the successful example of manual
feature is SIFT~\cite{lowe1999object}, which is widely used by many
computer vision algorithms.
In recent years, automatic feature engineering has drawn more
attentions. Compared to the manual feature
engineering, automatic feature engineering approaches allow to learn
useful features without strong expert knowledge about the data. Among
these learning approaches, deep neural networks are intensively studied.
The most common approach for deep feature learning is full
supervision, which usually requires large number of manual
annotations~\cite{yosinski2014transferable,zhou2016learning,wen2016discriminative}.
Unfortunately, such data is not always available for certain tasks or
needs to invest a huge amount of human labors to create.  In contrast,
unlabeled data is often plentiful (\ie cellphone photos in Instagram
or Youtube videos).

@@

So far, we have explained four important related computer vision
concepts. In the following section, we are going to discuss our main
contributions.
